### 2020-04-22 google_search.Rmd

---
title: "google_search"
author: "Federico Chung"
date: "4/21/2020"
output: html_document
---

```{r}
library(ggplot2)
library(dplyr)
```

```{r}
google_search<-read.csv("googlesearch.csv")

```

```{r}
google_search<-google_search%>%
  mutate(Day = paste(as.character(Day),"20",sep =""))%>%
  mutate(Day = as.Date(Day, format = "%m/%d/%Y"))
```

```{r}
ChinaVirus<-google_search%>%
  select(Day,contains("china.virus"))
KungFlu<-google_search%>%
  select(Day,contains("Kung.Flu"))
```

```{r}

states<- read.csv("us_states.csv")

states<-states%>%
  select(Alabama, AL)
states
colnames(states)<-(c("StateName","State"))

alabama<-c(NA,NA)

all_states<- rbind(states,alabama)

all_states

states_final_1<-all_states%>%
  mutate(StateName=as.character(StateName),State = as.character(State))%>%
  replace_na(list(StateName = "Alabama", State = "AL"))

states_final_2<-rbind(states_final_1,alabama)

states_final_2

states_final<-states_final_2%>%
  mutate(StateName=as.character(StateName),State = as.character(State))%>%
  replace_na(list(StateName = "District of Columbia", State = "DC"))%>%
  arrange(State)

states_final

colnames(ChinaVirus) <- c("Day",states_final$State)

Final_ChinaVirus<-ChinaVirus%>%
  gather("State","ChinaVirusInterest",2:ncol(ChinaVirus))

colnames(KungFlu) <- c("Day",states_final$State)

Final_KungFlu<-KungFlu%>%
  gather("State","KungFluInterest",2:ncol(KungFlu))

Time_series_Google<-Final_ChinaVirus%>%
  left_join(Final_KungFlu)

write.csv(Time_series_Google,"Time_series_Google.csv")
```

```{r}
day1<-as.Date("03/10/2020",format = "%m/%d/%Y")
day2<-as.Date("03/24/2020",format = "%m/%d/%Y")

Google_final<-Time_series_Google%>%
  filter(Day<=day2)%>%
  group_by(State)%>%
  summarize(mean_ChinaVirusInterest = mean(ChinaVirusInterest),
            mean_KungFluInterest = mean(KungFluInterest))

write.csv(Google_final,"Google_final.csv")
```