### 2020-04-22 merged data/merging_datasets.Rmd

---
title: "merging_datasets"
author: "Federico Chung"
date: "4/21/2020"
output: html_document
---

```{r}
library(ggplot2)
library(dplyr)
```

```{r}
demographic<-read.csv("Demographic_Data.csv")
covid<-read.csv("states_current.csv")
stayhome<-read.csv("quarantine.csv")
google<-read.csv("Google_final.csv")
```

```{r}
colnames(demographic)[2] <- "State"

covid<-covid%>%
  mutate(dateModified = as.Date(substr(dateModified,start = 0,stop = 10)),
         dateChecked = as.Date(substr(dateChecked,start = 0,stop = 10)))%>%
  select(-c(notes,hash))
colnames(covid)[1] <- "State"

stayhome<-stayhome %>% replace_with_na(replace = list(StayAtHome_date = "#N/A"))
no_na<-subset(stayhome,stayhome$StayAtHome_date!="NA") 
na<-stayhome%>%
    anti_join(no_na)%>%
    mutate(StayAtHome_date =as.Date(as.character(StayAtHome_date),format = "%m-%d-%Y"), na.rm = TRUE)
no_na1<-no_na%>%
  mutate(StayAtHome_date =as.Date(as.character(StayAtHome_date),format = "%m/%d/%Y"), na.rm = TRUE)
stayhomefinal<-rbind(no_na1,na)

all<-demographic%>%
  left_join(covid)%>%
  left_join(stayhomefinal)%>%
  left_join(google)

all

write.csv(all,"FullDataSet.csv")
```