### 2020-04-30 Checkpoint 4.Rmd
Update Checkpoint 4.Rmd
tables

---
title: "Checkpoint 4"
author: "Will Madairy, Sofia Pozsonyiova, Quinn Rafferty, Federico Chung"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
header-includes:
  - \usepackage{tabularx}
output:
  pdf_document: default
  html_document: default
---

```{r, warning=FALSE, include=FALSE}
library(dplyr)
library(usmap)
library(maps)
library(ggplot2)
library(tidyverse)
library(ggmap)
library(viridis)
library(rgdal)
library(gridExtra)
library(plotly)
require(janitor)
require(reshape2)
require(tidyr)
require(rstan)
require(rstanarm)
require(bayesplot)
```

```{r}
################################## To do before Alicia with your fave model ##########################################
# mcmc_trace()
# mcmc_dens_overlay()
# pp_check()

### For predicting one case 
          # new_308 <- data.frame(Days = 9, Subject =308)
          # 
          # set.seed(454)
          # pred_308 <- posterior_predict(
          #   model_4, 
          #   newdata = new_308)
          # 
          # pred_308 <- data.frame(y_new = pred_308[-1])
          # 
          # ggplot(pred_308, aes(x=y_new))+geom_density()
          # 
          # summary(pred_308)

## For correlation check (type like model 2 from activity 19)
          # Store the chains
          # model_1_df <- as.array(model_1) %>% 
          #   melt %>% 
          #   pivot_wider(names_from = parameters, values_from = value)
          # 
          # # Wrangle the chains
          # # Note the need for `...` around the sigma_sq_b
          # model_1_df <- model_1_df %>% 
          #   mutate(sigma_sq_w = `sigma`, sigma_sq_b = `Sigma[Region:(Intercept),(Intercept)]`) %>% 
          #   mutate(correlation = (sigma_sq_b/(sigma_sq_w+sigma_sq_b)))
          # 
          # ggplot(model_1_df, aes(x = correlation)) + 
          #   geom_density(alpha = 0.5)
```

```{r, include=FALSE}
prediction_summary_data <- function(y, yrep, prob_inner = 0.5, prob_outer = 0.95){
  # Calculate summary statistics of simulated 
  # posterior predictive models for each case
  l_outer <- function(x){quantile(x, (1-prob_outer) / 2)}
  l_inner <- function(x){quantile(x, (1-prob_inner) / 2)}
  u_inner <- function(x){quantile(x, 1 - (1-prob_inner) / 2)}
  u_outer <- function(x){quantile(x, 1 - (1-prob_outer) / 2)}
  df <- data.frame(yrep) %>% 
    summarize_all(list(mean, sd, median, mad, l_outer, l_inner, u_inner, u_outer)) %>%
    unlist() %>% 
    matrix(., length(y), 8) %>% 
    data.frame()
  names(df) <- c("post_mean", "post_sd", "post_median", "post_mad", "l_outer", "l_inner", "u_inner", "u_outer")
  data.frame(cbind(y, df))
}

prediction_summary <- function(y, yrep, prob_inner = 0.5, prob_outer = 0.95){
  # This function summarizes the predictions across all cases
  pred_data <- prediction_summary_data(y, yrep, prob_inner = prob_inner, prob_outer = prob_outer) %>% 
    mutate(error = y - post_median) %>% 
    mutate(error_scaled = error / post_mad) %>% 
    mutate(within_inner = (y >= l_inner) & (y <= u_inner)) %>% 
    mutate(within_outer = (y >= l_outer) & (y <= u_outer))

  pred_summary <- pred_data %>% 
    summarize(mae = median(abs(error)), 
      mae_scaled = median(abs(error_scaled)),
      within_inner = mean(within_inner),
      within_outer = mean(within_outer)
    )
  names(pred_summary)[3] <- paste0("within_", prob_inner*100)
  names(pred_summary)[4] <- paste0("within_", prob_outer*100)

  pred_summary
}
```

```{r, warning=FALSE}
Finaldata<-read_csv("time_series_final_Data.csv")
```

```{r} 
Finaldata <- Finaldata %>% mutate(Day= as.Date(Day))
day_initial<- as.Date("2020-03-14")
day_final<- as.Date("2020-03-21")
Finaldata<- Finaldata%>%
  filter(Day>=day_initial)%>%
  filter(Day<=day_final)
```

# Quinn plots stuff
```{r}
a<- ggplot(Finaldata, aes(x = ChinaVirusInterest, fill = as.factor(State))) + 
  geom_density(alpha = 0.5)+ ggtitle("China Virus Density by State")

b<- ggplot(Finaldata, aes(x = Day, y = ChinaVirusInterest, color = factor(State))) + 
    geom_point() + 
    geom_smooth(method = "lm", se = FALSE) + 
    facet_wrap(~ State) + 
    theme(legend.position = "right")+
    ggtitle("China Virus Interest by Day by State")

c<- ggplot(Finaldata, aes(x = ChinaVirusInterest, color = factor(StateColor))) + 
    geom_histogram() + 
    facet_wrap(~ Winner) +
    theme(legend.position = "right") + 
    scale_colour_manual( values=c("red"= "red3", "blue"= "royalblue", "purple"="darkorchid3")) +
    ggtitle("China Virus Google by State Poltical Party Winner")

d <-  ggplot(Finaldata, aes(y=positive, x=Winner, fill=(Winner)))+
      geom_boxplot() +
      facet_wrap(~Region)+
      theme(legend.position = "none") +
      scale_fill_manual(values = c("Democrat"="blue3", "Republican"= "red3"))+
      lims(y= c(0,1000))+
      ggtitle("Positive COVID-19 Cases by State Political Party Winner & Region")

ggplotly(a)
ggplotly(b)
ggplotly(c) #Shows that Democrat states tend to have a higher interest in Googling China Virus 
ggplotly(d)

#################################### Notes!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!

# What affects the China Interest search, is it demographic, policitical or COVID based? 
## From Alicia:: Start by thinking, take the four models from the sleep study and just think how we can go thorugh those four models in the context of our work.

## Chosen variables: Number of COVID-19 cases, State color, percent black, percent white, income
```

#Quinn making models
```{r, cahce=TRUE}
# 
# model_1 <- stan_glmer(
#   ChinaVirusInterest ~ (1 | State), #Gives us random intercepts
#   data = Finaldata, family = gaussian, chains = 4, iter = 5000*2
# )
```

```{r}
# head(data.frame(summary(model_1)),-2)
```

#Quinn making models
```{r, warning=FALSE, cache=TRUE}
# model_2 <- stan_glmer(
#   ChinaVirusInterest ~ Day + (Day | State),
#   data = Finaldata, family = gaussian,
#   chains = 4, iter = 100, refresh = 0
# )

```

```{r}
# head(data.frame(summary(model_2)),-2)
```

#Quinn makes models
```{r, warning = FALSE, cache=TRUE}
# model_3 <- stan_glmer(
#   ChinaVirusInterest ~ Day + StateColor + (Day | State),#Random intercepts and slopes when we put Day in
#   data = Finaldata, family = gaussian,
#   chains = 4, iter = 100, refresh = 0
# )
```

```{r}
# head(data.frame(summary(model_3)),-2)
```

#Quinn makes models
```{r, warning = FALSE, cache=TRUE}
# model_4 <- stan_glmer(
#   ChinaVirusInterest ~ Day + totalTestResults + (Day | State),#Random intercepts and slopes when we put Day in
#   data = Finaldata, family = gaussian,
#   chains = 4, iter = 100, refresh = 0
# )
```

```{r}
# head(data.frame(summary(model_4)),-2)
```

```{r, warning = FALSE, cache=TRUE}
model_100 <- stan_glmer(
  ChinaVirusInterest ~ Day + totalTestResults + positive + percent_white +  (Day | State),#Random intercepts and slopes when we put Day in
  data = Finaldata, family = gaussian,
  chains = 4, iter = 5000*2, refresh = 0
)

```

```{r}
head(data.frame(summary(model_100)),-2)
```
```{r}
mcmc_trace(model_100)
mcmc_dens_overlay(model_100) 
```

```{r}
pp_check(model_100)
```

```{r}
set.seed(454)
pred_1 <- posterior_predict(
  model_100,
  newdata = Finaldata, transform = TRUE)

# pred_5 <- posterior_predict(
#   model_5,
#   newdata = Finaldata, transform = TRUE)

prediction_summary(y = Finaldata$ChinaVirusInterest,
  yrep = pred_1)

# prediction_summary(y = Finaldata$ChinaVirusInterest,
#   yrep = pred_5)
```

```{r}
### For predicting one case 
new_CA <- Finaldata%>% 
          filter(State=="CA") #filter the row of interest

set.seed(454)
pred_CA <- posterior_predict(
  model_100,
  newdata = new_CA)

pred_2 <- data.frame(y_new = pred_CA[-1])

ggplot(pred_2, aes(x=y_new))+geom_density()

summary(pred_2)

```

```{r}
# For correlation check (type like model 2 from activity 19)
# Store the chains
model_2_df <- as.array(model_2) %>%
  melt %>%
  pivot_wider(names_from = parameters, values_from = value)

# Wrangle the chains
# Note the need for `...` around the sigma_sq_b
model_2_df <- model_2_df %>%
  mutate(sigma_sq_w = `sigma`, sigma_sq_b = `Sigma[State:(Intercept),(Intercept)]`) %>%
  mutate(correlation = (sigma_sq_b/(sigma_sq_w+sigma_sq_b)))

ggplot(model_2_df, aes(x = correlation)) +
  geom_density(alpha = 0.5)
```

#Sofia's Models:

##Model 5:

```{r, warning = FALSE, cache=TRUE}
set.seed(454)
Finaldata<-
  Finaldata%>%
  mutate(Day=as.numeric(Day))

model_5 <- stan_glmer(
  ChinaVirusInterest ~ Day + StateColor + (Day | State), #Gives us random intercepts
  data = Finaldata, family = gaussian, chains = 4, iter = 5000*2
)
```

```{r}
head(data.frame(summary(model_5)),-2)
```

Checks:

```{r}
# Trace plots
#mcmc_trace(model_5)

# Density plots
#mcmc_dens_overlay(model_5)
```

```{r}
# y = original sample, yrep = MCMC simulation
set.seed(454)
pp_check(model_5, nreps = 50)
```

```{r}
# Store the chains
model_5_df <- as.array(model_5) %>%
  melt %>%
  pivot_wider(names_from = parameters, values_from = value)

# Wrangle the chains
# Note the need for `...` around the sigma_sq_b
model_5_df <- model_5_df %>%
  mutate(sigma_sq_w = `sigma`, sigma_sq_b = `Sigma[State:(Intercept),(Intercept)]`) %>%
  mutate(correlation = (sigma_sq_b/(sigma_sq_b+sigma_sq_w)))

ggplot(model_5_df, aes(x = correlation)) +
  geom_density(alpha = 0.5)
```

```{r}
#set.seed(454)
my_data <- data.frame(StateColor = "blue", State = "FL", Day = 18336)
# Simulate a set of predictions for each case in newdata
#my_prediction <- posterior_predict(
 # model_5, 
  #newdata = my_data)

#my_prediction <- data.frame(y_new = my_prediction[,1])

#ggplot(my_prediction, aes(x = y_new)) + 
 # geom_density()

#summary(my_prediction)
```

##Model 9:

StateColor +percent_asian + totalTestResults + (StateColor | State)

```{r, warning = FALSE, cache=TRUE}
set.seed(454)
Finaldata<-
  Finaldata%>%
  mutate(Day=as.numeric(Day))

model_9 <- stan_glmer(
  ChinaVirusInterest ~ StateColor +percent_asian + totalTestResults + (StateColor | State), #Gives us random intercepts
  data = Finaldata, family = gaussian, chains = 4, iter = 1000*2
)
```

```{r}
head(data.frame(summary(model_9)),-2)
```

```{r}
# y = original sample, yrep = MCMC simulation
set.seed(454)
pp_check(model_9, nreps = 50)
```

```{r}
# Store the chains
model_9_df <- as.array(model_9) %>% 
  melt %>% 
  pivot_wider(names_from = parameters, values_from = value)

# Wrangle the chains
# Note the need for `...` around the sigma_sq_b
model_9_df <- model_9_df %>% 
  mutate(sigma_sq_w = `sigma`, sigma_sq_b = `Sigma[State:(Intercept),(Intercept)]`) %>% 
  mutate(correlation = (sigma_sq_b/(sigma_sq_b+sigma_sq_w)))

ggplot(model_9_df, aes(x = correlation)) + 
  geom_density(alpha = 0.5)

```

```{r}
#set.seed(454)
my_data <- data.frame(StateColor = "red", State = "FL")

# Simulate a set of predictions for each case in newdata
my_prediction <- posterior_predict(
  model_9,
   newdata = my_data)

my_prediction <- data.frame(y_new = my_prediction[,1])

ggplot(my_prediction, aes(x = y_new)) +
 geom_density()

summary(my_prediction)
```

ChinaVirusInterest ~ Day + StateColor + (Day | State)	mean	mcse	sd	X10.	X50.	X90.	n_eff	Rhat
	(Intercept)	-1.94E+04	4.20E+01	6.48E+03	-2.77E+04	-1.94E+04	-1.11E+04	23821
	Day	1.06E+00	2.29E-03	3.54E-01	6.06E-01	1.06E+00	1.51E+00	23821
	StateColorpurple	-4.47E-01	3.93E-02	3.96E+00	-5.46E+00	-4.45E-01	4.57E+00	10177
	StateColorred	-4.09E+00	3.24E-02	3.35E+00	-8.39E+00	-4.09E+00	1.79E-01	10690

ChinaVirusInterest ~ StateColor + percent_asian + (Day | State)								
	(Intercept)	4.36E+01	3.80E-02	3.16E+00	3.96E+01	4.35E+01	4.76E+01	6895
	StateColorpurple	-8.65E-01	5.00E-02	4.23E+00	-6.28E+00	-8.61E-01	4.40E+00	7182
	StateColorred	-4.48E+00	4.29E-02	3.68E+00	-9.18E+00	-4.50E+00	2.35E-01	7384
	percent_asian	-8.57E+00	3.33E-01	3.10E+01	-4.82E+01	-8.11E+00	3.03E+01	8710

ChinaVirusInterest ~ StateColor +percent_asian + totalTestResults + (StateColor | State)								
	(Intercept)	4.27E+01	7.19E-02	3.00E+00	3.90E+01	4.27E+01	4.65E+01	1741
	StateColorpurple	-2.52E-01	9.18E-02	4.05E+00	-5.38E+00	-2.01E-01	4.85E+00	1949
	StateColorred	-3.82E+00	9.17E-02	3.77E+00	-8.63E+00	-3.87E+00	8.40E-01	1685
	percent_asian	-1.07E+01	5.48E-01	2.86E+01	-4.61E+01	-1.05E+01	2.60E+01	2720
	totalTestResults	3.17E-04	4.75E-06	2.78E-04	-4.82E-05	3.19E-04	6.61E-04	3409

#Fed's visualizations:
```{r}
#for visuals
install.packages("ggforce")
library(ggforce)
ggplot(Finaldata, aes(x = as.numeric(Day), y = ChinaVirusInterest, color = factor(State))) + 
    geom_point() + 
    geom_smooth(method = "lm", se = FALSE) + 
    facet_wrap_paginate(~ State, ncol = 3, nrow = 3, page = 1) + 
    theme(legend.position = "none")
```

```{r}
ggplot(Finaldata, aes(x = as.numeric(Day), y = ChinaVirusInterest, color = factor(State))) + 
    geom_point() + 
    geom_smooth(method = "lm", se = FALSE) + 
    facet_wrap_paginate(~ State, ncol = 3, nrow = 3, page = 2) + 
    theme(legend.position = "none")
```

```{r}
ggplot(Finaldata, aes(x = as.numeric(Day), y = ChinaVirusInterest, color = factor(State))) + 
    geom_point() + 
    geom_smooth(method = "lm", se = FALSE) + 
    facet_wrap_paginate(~ State, ncol = 3, nrow = 3, page = 3) + 
    theme(legend.position = "none")
```

```{r}
ggplot(Finaldata, aes(x = as.numeric(Day), y = ChinaVirusInterest, color = factor(State))) + 
    geom_point() + 
    geom_smooth(method = "lm", se = FALSE) + 
    facet_wrap_paginate(~ State, ncol = 3, nrow = 3, page = 4) + 
    theme(legend.position = "none")
```

```{r}
ggplot(Finaldata, aes(x = as.numeric(Day), y = ChinaVirusInterest, color = factor(State))) + 
    geom_point() + 
    geom_smooth(method = "lm", se = FALSE) + 
    facet_wrap_paginate(~ State, ncol = 3, nrow = 3, page = 5) + 
    theme(legend.position = "none")
```

```{r}
ggplot(Finaldata, aes(x = as.numeric(Day), y = ChinaVirusInterest, color = factor(State))) + 
    geom_point() + 
    geom_smooth(method = "lm", se = FALSE) + 
    facet_wrap_paginate(~ State, ncol = 3, nrow = 3, page = 6) + 
    theme(legend.position = "none")
```

```{r}
Finaldata%>%
  mutate(perc_population_positive = positive/total_population)%>%
  group_by(State)%>%
  mutate(mean_corona_interest = mean(ChinaVirusInterest),
         max = max(ChinaVirusInterest),
         median = median(ChinaVirusInterest))%>%
  ggplot(aes(x = mean_corona_interest, y = perc_population_positive, color = percent_asian))+
  geom_point()+
  scale_colour_gradient2(low = "blue", mid = "lightblue", high = "red")
```

#Fed's Models:

Variables of interest:
- % white
- postive rate (out of the tested)

```{r}
glimpse(Finaldata)
model_data<-Finaldata%>%
  mutate(Date = as.numeric(Day))%>%
  mutate(positive_rate = positive/ totalTestResults,
         ChinaVirusInterest_dec = ChinaVirusInterest/100)%>%
  select(State, ChinaVirusInterest, Day, percent_white,positive_rate,ChinaVirusInterest_dec)
model_data<- na.omit(model_data)

colSums(is.na(model_data))
```
```{r}
model_10 <- stan_glmer(
  ChinaVirusInterest ~ Day + (1 | State) + percent_white,
  data = model_data, family = gaussian, chains = 4, iter = 1000
)
```
```{r}
head(data.frame(summary(model_10)),-2)
```

```{r}
# Store the chains
model_10_df <- as.array(model_10) %>% 
  melt %>% 
  pivot_wider(names_from = parameters, values_from = value)

# Wrangle the chains
# Note the need for `...` around the sigma_sq_b
model_10_df <- model_10_df %>% 
  mutate(sigma_sq_w = `sigma`, sigma_sq_b = `Sigma[State:(Intercept),(Intercept)]`) %>% 
  mutate(correlation = (sigma_sq_b/(sigma_sq_w+sigma_sq_b)))

ggplot(model_10_df, aes(x = correlation)) + 
  geom_density(alpha = 0.5)
```
```{r}
set.seed(454)
#random intercept and random slope, percent white
model_11 <- stan_glmer(
  ChinaVirusInterest ~ (Day | State) + percent_white,
  data = model_data, family = gaussian, chains = 4, iter = 3000
)
```
```{r}
head(data.frame(summary(model_11)),-2)
```

```{r}
# Store the chains
model_11_df <- as.array(model_11) %>% 
  melt %>% 
  pivot_wider(names_from = parameters, values_from = value)

# Wrangle the chains
# Note the need for `...` around the sigma_sq_b
model_11_df <- model_11_df %>% 
  mutate(sigma_sq_w = `sigma`, sigma_sq_b = `Sigma[State:(Intercept),(Intercept)]`) %>% 
  mutate(correlation = (sigma_sq_b/(sigma_sq_w+sigma_sq_b)))

ggplot(model_10_df, aes(x = correlation)) + 
  geom_density(alpha = 0.5)
```

```{r}
set.seed(454)

model_12<- stan_glmer(
  ChinaVirusInterest ~ Day + (1 | State) + percent_white +positive_rate,
  data = model_data, family = gaussian, chains = 4, iter = 3000
)
```
```{r}
head(data.frame(summary(model_12)),-2)
```

```{r}
# Store the chains
model_12_df <- as.array(model_12) %>% 
  melt %>% 
  pivot_wider(names_from = parameters, values_from = value)

# Wrangle the chains
# Note the need for `...` around the sigma_sq_b
model_12_df <- model_12_df %>% 
  mutate(sigma_sq_w = `sigma`, sigma_sq_b = `Sigma[State:(Intercept),(Intercept)]`) %>% 
  mutate(correlation = (sigma_sq_b/(sigma_sq_w+sigma_sq_b)))

ggplot(model_12_df, aes(x = correlation)) + 
  geom_density(alpha = 0.5)
```

```{r}
model_data<-model_data%>%
  select(-c(ChinaVirusInterest))
set.seed(454)
#random intercept and random slope, percent white and positive rate = positive/ total tests
#iterations might be low need to sample up if we want to make the models better
model_13 <- stan_glmer(
  ChinaVirusInterest_dec ~ (Day | State) + percent_white + positive_rate,
  data = model_data, family = gaussian, chains = 4, iter = 3000
)
```
```{r}
head(data.frame(summary(model_13)),-2)
```

```{r}
pp_check(model_13)
```

```{r}
# Store the chains
model_13_df <- as.array(model_13) %>% 
  melt %>% 
  pivot_wider(names_from = parameters, values_from = value)

# Wrangle the chains
# Note the need for `...` around the sigma_sq_b
model_13_df <- model_13_df %>% 
  mutate(sigma_sq_w = `sigma`, sigma_sq_b = `Sigma[State:(Intercept),(Intercept)]`) %>% 
  mutate(correlation = (sigma_sq_b/(sigma_sq_w+sigma_sq_b)))

ggplot(model_13_df, aes(x = correlation)) + 
  geom_density(alpha = 0.5)
```
```{r}
set.seed(454)
pred_13 <- posterior_predict(
  model_13,
  newdata = model_data, transform = TRUE)

prediction_summary(y = model_data$ChinaVirusInterest,
  yrep = pred_13)
```
#Final model Fed
```{r}
### For predicting one case

new_FL <- data.frame(Day = 7 , State = "FL",percent_white = 0.5, positive_rate = 0.11) #Doesn't recognize state
set.seed(454)
pred_FL <- posterior_predict(
  model_13,
  newdata = new_FL)

pred_FL <- data.frame(y_new = pred_FL[-1])

ggplot(pred_FL, aes(x=y_new))+geom_density()

summary(pred_FL)
```

# Will's Models

-Variables of Interest:
*Pct Black
*Pct Positive
*Pct Negative

```{r}
mod_data<-Finaldata%>%
  mutate(Date = as.numeric(Day))%>%
  mutate(positive_rate = positive/ totalTestResults)%>%
  mutate(negative_rate = 1-positive_rate) %>% 
  select(State, ChinaVirusInterest, Day, percent_white, percent_black, positive_rate, negative_rate, KungFluInterest, Region, StateColor)
model_data<- na.omit(model_data)
```

Model 14
```{r}
set.seed(454)
#Neg Rate
model_14 <- stan_glmer(
  ChinaVirusInterest ~ (1 | State) + negative_rate,
  data = mod_data, family = gaussian, chains = 4, iter = 1000
)
```

```{r}
head(data.frame(summary(model_14)),-2)
```

Checking Model 14
```{r}
mcmc_trace(model_14)
mcmc_dens(model_14)
```
```{r}
pp_check(model_14, nreps=50)
```

```{r}
# Store the chains
model_14_df <- as.array(model_14) %>% 
  melt %>% 
  pivot_wider(names_from = parameters, values_from = value)

# Wrangle the chains
# Note the need for `...` around the sigma_sq_b
model_14_df <- model_14_df %>% 
  mutate(sigma_sq_w = `sigma`, sigma_sq_b = `Sigma[State:(Intercept),(Intercept)]`) %>% 
  mutate(correlation = (sigma_sq_b/(sigma_sq_w+sigma_sq_b)))

ggplot(model_14_df, aes(x = correlation)) + 
  geom_density(alpha = 0.5)
```

```{r}
set.seed(454)
my_data <- data.frame(negative_rate = .9, State = "IL", percent_black = .12)

 # Simulate a set of predictions for each case in newdata
my_prediction <- posterior_predict(
   model_14,
   newdata = my_data)

my_prediction <- data.frame(y_new = my_prediction[,1])

ggplot(my_prediction, aes(x = y_new)) +
  geom_density()

summary(my_prediction)
```

```{r}
model_15<-stan_glmer(
  ChinaVirusInterest ~ (Day|State) + negative_rate,
  data=mod_data, family=gaussian, chains =4, iter = 5000
)
```

```{r}
head(data.frame(summary(model_15)),-2)
```

```{r}
# Store the chains
model_15_df <- as.array(model_15) %>%
  melt %>%
  pivot_wider(names_from = parameters, values_from = value)

# Wrangle the chains
# Note the need for `...` around the sigma_sq_b
model_15_df <- model_15_df %>%
  mutate(sigma_sq_w = `sigma`, sigma_sq_b = `Sigma[State:(Intercept),(Intercept)]`) %>%
  mutate(correlation = (sigma_sq_b/(sigma_sq_w+sigma_sq_b)))

ggplot(model_15_df, aes(x = correlation)) +
  geom_density(alpha = 0.5)
```

```{r}
model_16<-stan_glmer(
  ChinaVirusInterest~ (1|State) + percent_black + negative_rate, 
  data = mod_data, family = gaussian, chains = 4, iter = 5000
)
```

```{r}
head(data.frame(summary(model_16)),-2)
```

```{r}
mcmc_trace(model_16)
mcmc_dens(model_16)
```

```{r}
pp_check(model_16, iter=50)
```

```{r}
# Store the chains
model_16_df <- as.array(model_16) %>% 
  melt %>% 
  pivot_wider(names_from = parameters, values_from = value)

# Wrangle the chains
# Note the need for `...` around the sigma_sq_b
model_16_df <- model_16_df %>% 
  mutate(sigma_sq_w = `sigma`, sigma_sq_b = `Sigma[State:(Intercept),(Intercept)]`) %>% 
  mutate(correlation = (sigma_sq_b/(sigma_sq_w+sigma_sq_b)))

ggplot(model_16_df, aes(x = correlation)) + 
  geom_density(alpha = 0.5)

```

```{r}
set.seed(454)
my_data <- data.frame(negative_rate = .9, State = "IL", percent_black = .12)

 # Simulate a set of predictions for each case in newdata
my_prediction <- posterior_predict(
   model_16,
   newdata = my_data)

my_prediction <- data.frame(y_new = my_prediction[,1])

ggplot(my_prediction, aes(x = y_new)) +
  geom_density()

summary(my_prediction)
```

```{r}
# # Pct black, Neg Rate: Random effects
# model_17<-stan_glmer(
#   ChinaVirusInterest~ (Day|State) + percent_black + negative_rate, 
#   data = mod_data, family = gaussian, chains = 4, iter = 2*5000
# 
# )
```

```{r}
# head(data.frame(summary(model_17)),-2)
```

```{r}
# model_17_df <- as.array(model_17) %>% 
#   melt %>% 
#   pivot_wider(names_from = parameters, values_from = value)
# 
# # Wrangle the chains
# # Note the need for `...` around the sigma_sq_b
# model_17_df <- model_17_df %>% 
#   mutate(sigma_sq_w = `sigma`, sigma_sq_b = `Sigma[State:(Intercept),(Intercept)]`) %>% 
#   mutate(correlation = (sigma_sq_b/(sigma_sq_w+sigma_sq_b)))
# 
# ggplot(model_17_df, aes(x = correlation)) + 
#   geom_density(alpha = 0.5)
```

Model Outputs:

Model 14: 
ChinaVirusInterest~ (1|State) + negative_rate

mean	mcse	sd	X10.	X50.	X90.	n_eff	Rhat

44.38040197	0.11616343	4.2066034	39.0747590	44.36959567	49.6577069	1311	1.0010474
-4.03732266	0.13077873	4.8298203	-10.2151294	-4.12905314	2.2624868	1364	1.0007950

Model 16: 
ChinaVirusInterest~ (1|State) + percent_black + negative_rate

mean	mcse	sd	X10.	X50.	X90.	n_eff	Rhat

(intercept) 45.38818892	0.059114913	4.851982	39.1598750	45.42176405	51.6208513	6737	1.0001366
percent black -4.81756453	0.218542483	14.099591	-22.6838613	-5.02290077	13.1161595	4162	1.0008973
negative_rate -4.48821663	0.053252003	4.937795	-10.7851047	-4.47175240	1.7764630	8598	1.0002747